taxadb: A High-Performance Local Taxonomic Database Interface

### 2020-02-12 paper/manuscript.Rmd
Final edits

journal: "Methods in Ecology & Evolution"
date: "`r Sys.Date()`"
bibliography: refs.bib
layout: 3p
output: 
  rticles::elsevier_article:
    includes:
      in_header: preamble.tex
---

@@

<!--
### Installation

`taxadb` can be installed from GitHub using the `remotes` R package:

```{r eval=FALSE}
remotes::install_github("ropensci/taxadb")
```
-->

@@

If we first resolve names used in each data set into shared identifiers, (for instance, using the Catalogue of Life), we discover that there is far more overlap in the species coverage than we might have realized at first.  First, we just add an ID column to each table by looking up the Catalog of Life identifier for the name provided:

@@

`get_ids` offers a first pass at matching scientific names to id, but names may remain unresolved for a number of reasons. First, a name may match to multiple accepted names, as in the case of a species that has been split. By design, these cases are left to be resolved by the researcher using the 'filter_' functions to filter underlying taxonomic tables for additional information. A name may also be unresolved due to typos or improper formatting. `clean_names` addresses common formatting issues such as the inclusion of missing species epithets (e.g. `Accipiter sp.`) that prevent matches to the Genus, or intraspecific epithets such as `Colaptes auratus cafer` that prevent matches to the binomial name. These modifications are not appropriate in all settings and should be used with care. Spell check of input names is outside the scope of `taxadb`.

@@

We see that *Zosterops rendovae* matches two accepted names which the user will have to choose between (Table 6).
This is an example of how `taxadb` seeks to provide users with information from existing authorities and names providers,
rather than make a potentially arbitrary decision.  Because they return `data.frame`s,  `filter_*` functions provide both potential matches.  Note that the simpler `get_` functions (`get_ids()`) consider multiple name as `NA` for the `id`, making them suitable for automated pipelines where manual resolution of duplicates is not an option.

>>>>

journal: "Methods in Ecology & Evolution"
date: "`r Sys.Date()`"
bibliography: refs.bib
layout: 3p
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output: 
  rticles::elsevier_article:
    includes:
      in_header: preamble.tex
---

@@

@@

If we first resolve names used in each data set into shared identifiers, (for instance, using the Catalogue of Life), we discover that there is far more overlap in the species coverage than we might have initially realized. First, we just add an ID column to each table by looking up the Catalog of Life identifier for the name provided:

@@

`get_ids` offers a first pass at matching scientific names to id, but names may remain unresolved for a number of reasons. First, a name may match to multiple accepted names, as in the case of a species that has been split. By design, these cases are left to be resolved by the researcher using the `filter_` functions to filter underlying taxonomic tables for additional information. A name may also be unresolved due to typos or improper formatting. `clean_names` addresses common formatting issues such as the inclusion of missing species epithets (e.g. `Accipiter sp.`) that prevent matches to the Genus, or intraspecific epithets such as `Colaptes auratus cafer` that prevent matches to the binomial name. These modifications are not appropriate in all settings and should be used with care. Spell check of input names is outside the scope of `taxadb`.

@@

We see that *Zosterops rendovae* matches two accepted names which the user will have to choose between (Table 6).
This is an example of how `taxadb` seeks to provide users with information from existing authorities and names providers,
rather than make a potentially arbitrary decision.  Because they return `data.frame`s,  `filter_` functions provide both potential matches.  Note that the simpler `get_` functions (`get_ids()`) consider multiple name matches as `NA` for the `id`, making them suitable for automated pipelines where manual resolution of duplicates is not an option.
